---
title: "Letters"
author: "Kevin Cummiskey"
date: "8/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

cadets <- read_csv(file = "https://raw.githubusercontent.com/kfcaby/MA376AY21/master/data/memorizing%20letters%20(Exploration%201.1)")
```

## Step 3

First, we'll investigate the response variable.

```{r}
cadets %>% 
  ggplot(aes(x = factor(group), y = letters)) +
  geom_boxplot()
```

The single mean model is:

$$y_i = \mu + \epsilon_i$$
$$\epsilon_i$$ has a Normal distribution with mean 0 and variance $$\sigma^2$$.

The separate means model is:

$$y_i = \mu_j + \epsilon_{ij}$$

```{r}
cadets %>% 
  group_by(group) %>% 
  summarize(mean = mean(letters))
```

We want to know the standard error of the single mean and separate means model.

Single mean model

```{r}
# calculate residuals
cadets = cadets %>% 
  mutate(single.mean = mean(letters),
         single.mean.residual = letters - single.mean)

cadets %>% 
  summarize(single.mean.se = sqrt(sum(single.mean.residual^2)/14))
```

Separate mean model

```{r}
# calculate residuals
cadets = cadets %>%
  group_by(group) %>% 
  mutate(group.mean = mean(letters))
  
cadets = cadets %>% 
  mutate(group.mean.residual = letters - group.mean)
 
#standard error 
cadets %>% 
  group_by() %>% 
  summarize(group.mean.se = sqrt(sum(group.mean.residual^2)/13))
```

Calculate Sums of squares total (SST)

```{r}
#distance from data point to the overall mean
# single mean residual is y_i - ybar

SST = cadets %>% 
  group_by() %>% 
  summarize(sum(single.mean.residual^2))
SST  

```


Calculate Sums of Squares Model (SSM)

```{r}
SSM = cadets %>% 
  group_by() %>% 
  summarize(sum((group.mean - single.mean)^2))
SSM
```

How would I do this in practice?

```{r}
model_group = cadets %>% 
  lm(letters ~ factor(group), data = .)

model_group %>% summary()

anova(model_group)

```

Null: there is no association between experimental group and number of letters memorized.
Alternative: there is an association between experimental group and number of letters memorized.

Calculate a $p$-value using simulation-based inference

```{r}
cadets.sim <- cadets

results <- c()

for(i in 1:1000){
  #shuffle the groups
  cadets.sim <- cadets.sim %>% 
    mutate(group.sim = sample(group))
  
  #calculate the means
  diff <- cadets.sim %>% 
    group_by(group.sim) %>% 
    summarize(group.mean = mean(letters)) %>% 
    mutate(diff = group.mean - lag(group.mean)) %>% 
    pull(diff)
  
  results[i] = diff[2]
}

sum(results < -10.167)/1000
```

Calculate a $p$-value using theory-based test.

```{r}
#t.test
cadets %>% 
  t.test(letters ~ group, data = .)

# regression notation
cadets %>% 
  lm(letters ~ group, data = .) %>% 
  summary()

```











```{r}
cadets.sim <- cadets

diffs <- c()

for(i in 1:5000){
  results <- cadets.sim %>% 
    mutate(group.sim = sample(group)) %>% 
    group_by(group.sim) %>% 
    summarize(group.mean = mean(letters)) %>% 
    mutate(diff = group.mean - lag(group.mean)) %>% 
    pull(diff)
  diffs[i] = results[2]
}
```

